Incorporating Gyroscope Information in Gesture Recognition

### 2011-12-17 paper/paper.tex

\paragraph{HMM}
Although we found a few examples where the cluster-matching algorithm
benefits from using a Hidden Markov Model, in most real-world examples
there is little benefit.  When running on all of the training data,
we

\begin{center}
\begin{tabular}{ | l | l | l | l | l | }
  \hline
  Normalization & Data & Features & Gesture Types & Accuracy \\ \hline
  No	&	Simple	&	Gyro	&	circle \& triangle	&	87.7\% \\ \hline
  No	&	Simple	&	Accel	&	circle \& triangle	&	87.7\% \\ \hline
  No	&	Simple	&	Accel and Gyro	&	circle \& triangle	&	96.7\% \\ \hline
  Yes	&	Simple	&	Accel and Gyro	&	circle \& triangle	&	90.7\% \\ \hline
  No	&	Complex	&	Accel and Gyro	&	circle \& triangle	&	89.5\% \\ \hline
  Yes	&	Complex	&	Accel and Gyro	&	circle \& triangle	&	91.9\% \\ \hline
  Yes	&	Complex	&	Accel	&	up-flip \& right-flip	&	91.0\% \\ \hline
  Yes	&	Complex	&	Gyro	&	up-flip \& right-flip	&	100\% \\ \hline
  Yes	&	Complex	&	Accel and Gyro	&	up-flip \& right-flip	&	100\% \\ \hline
  Yes	&	Complex	&	Accel	&	up-flip, right-flip, circle, triangle, and bowl	&	87.9\% \\ \hline
  Yes	&	Complex	&	Gyro	&	up-flip, right-flip, circle, triangle, and bowl	&	89.8\% \\ \hline
  Yes	&	Complex	&	Accel and Gyro	&	up-flip, right-flip, circle, triangle, and bowl	&	90.0\% \\
  \hline
\end{tabular}
\end{center}

>>>>

\paragraph{HMM}
Although we found a few examples where the cluster-matching algorithm
benefits from using a Hidden Markov Model, in most real-world examples
there is little benefit.  When running on all of the training data,
we found that adding the HMM actually decreased accuracy slightly,
from 96.7\% to 94.6\%, which may be due to random fluctuations.

However, when comparing the data for circles to our contrived data
set consisting of the first and second halves of each circle switched,
the using an HMM with alphabet size of 20 improved the results from chance
to 88.8\%.  Unfortunately we could not get such dramatic improvements with
any non contrived data.

\begin{center}
\begin{tabular}{ | l | l | l | l | l | }
  \hline
  Normalization & Features & Algorithm & Gestures Types & Accuracy \\ \hline
  Yes	&	Both  &	All & Cluster-matching with HMM (10) &	94.6\% \\ \hline
  Yes	&	Both	&	All & Original  &	89.6\% \\ \hline
  Yes	&	Accel &	All & Cluster-matching with HMM (10) &	92.0\% \\ \hline
  Yes	&	Accel &	All &	Original & 84.2\% \\ \hline
  Yes	&	Gyro  &	All & Cluster-matching with HMM (10) & 92.3\% \\ \hline
  Yes	&	Gyro  &	All & Original &  89.7\% \\ \hline
  Yes	&	Both	&	All &	Cluster-matching (10) &  96.7\% \\ \hline
  Yes	&	Both	&	All &	Cluster-matching (20) &  97.9\% \\ \hline
  Yes	&	Both  &	All &	Cluster-matching (40) & 	98.4\% \\ \hline
  No	&	Both	&	All &	Cluster-matching (40) & 	98.4\% \\ \hline
  Yes	&	Both	& Squares &	Cluster-matching (10) & 	100\% \\ \hline
  Yes	&	Both	& Circles and Reversed-circles  &	Cluster-matching (10) & 	Chance \\ \hline
  Yes	&	Both	& Circles and Reversed-circles  &	Cluster-matching with HMM (10) &	72.0\% \\ \hline
  Yes	&	Both  & Circles and Reversed-circles  &	Cluster-matching with HMM (20) & 88.8\% \\
  \hline
\end{tabular}
\end{center}
